Classical Simulation of Quantum Error Correction in a Fibonacci Anyon Code

### 2015-05-06 fib2.tex
Added Refs. Addressed Simon's comment about decoder efficiency.

Topologically ordered quantum systems in two dimensions offer tremendous promise for 
long-term storage and processing of quantum information~\cite{Brown2015, Terhal2015}. 
The topological features of such systems are insensitive to local 
perturbations~\cite{Bravyi2010, Bravyi2011a, Michalakis2013}, and have excitations 
displaying anyonic statistics~\cite{Wilczek1990}. These anyons can in general be used as 
quantum memories~\cite{Kitaev2003, Dennis2002} or to perform universal topological 
quantum computation~\cite{Freedman2002, Nayak2008}.

Quantum error correction is vital to harnessing the computational power of these topologically 
ordered systems. When coupled to a heat bath at any finite temperature, thermal fluctuations 
will create spurious anyons that diffuse and quickly corrupt the stored quantum 
information~\cite{Pastawski2010}. Thus, the passive protection provided by the mass gap 
at low temperature must be augmented by an \emph{active} procedure.

Almost all of the sizable research effort on active quantum error correction for topological 
systems has focused on the case of abelian anyons~\cite{Terhal2015} \cggb{revise citations here: poulin decoding stuff~\cite{Duclos-Cianci2010, Duclos-Cianci2010a, Duclos-Cianci2013}, what else?} \stf{Dan Browne's recent work, Bravyi's recent work, Austin Fowler? Do we really want to cite everyone that's ever decoded the toric code?}. These systems are 
well suited to studying quantum error correction because the dynamics of excitations in 
abelian anyon models (at the RG fixed point) can be simulated numerically very efficiently, allowing lattice 
simulations of decoding with over 1 million sites~\cite{Duclos-Cianci2010}. Standard algorithms to efficiently simulate these systems rely on the abelian nature of these particles, and in particular the fact that abelian anyons cannot be used for quantum computation.

@@

Here we consider a full simulation of quantum error correction in a two-dimensional lattice 
system with Fibonacci anyons, a class of nonabelian anyons that are universal for quantum 
computation~\cite{Freedman2002, Nayak2008}. Fibonacci anyons are experimentally motivated as the 
expected excitations of the $\nu=\frac{12}{5}$ fractional quantum Hall 
states~\cite{Slingerland2001}, and can also be realized in several spin 
models~\cite{Levin2005, Kapit2013, Palumbo2014} and composite 
heterostructures~\cite{Mong2014}.

\stf{cite:~\cite{Bonesteel2012}} \cggb{what is this citation for? something about experimentally relevant in engineered systems?}

@@

\dude{we should stress the distinction between
simulation the fib dynamics and the simulation of
the decoder (which is poly by design).. this is something
that doherty completely missed at first.. the decoder is
classical and it does not care what anyon system it is
looking at.}

@@

\stf{Is there a possibility of using MPS like Rob Pfeiffer to improve the simulations or 
include interactions?}
\dude{Our simulation is effectively doing a contraction of a 2d peps-like network. So
yeah, maybe there are connections.}

@@

\cggb {People we chatted to.} 
This work was supported by the ARC via EQuS project number CE11001013, by the US Army Research Office grant numbers W911NF-14-1-0098 and W911NF-14-1-0103, ERC grant QFTCMPS and by the cluster of excellence EXC 201 Quantum Engineering and Space-Time Research. STF also acknowledges support from an ARC Future Fellowship FT130101744.

>>>>

Topologically ordered quantum systems in two dimensions offer tremendous promise for 
long-term storage and processing of quantum information~\cite{Kitaev2003, Dennis2002, Nayak2008}. 
The topological features of such systems are insensitive to local 
perturbations~\cite{Bravyi2010, *Bravyi2011a, *Michalakis2013}, and have excitations 
displaying anyonic statistics~\cite{Wilczek1990}. These anyons can in general be used as 
quantum memories~\cite{Kitaev2003, Dennis2002} or to perform universal topological 
quantum computation~\cite{Freedman2002, Nayak2008}.

Quantum error correction is vital to harnessing the computational power of these topologically 
ordered systems. When coupled to a heat bath at any finite temperature, thermal fluctuations 
will create spurious anyons that diffuse and quickly corrupt the stored quantum 
information~\cite{Pastawski2010}. Thus, the passive protection provided by the mass gap 
at low temperature must be augmented by an \emph{active} decoding procedure.

Almost all of the sizable research effort on active quantum error correction for topological 
systems has focused on the case of abelian anyons~\cite{Dennis2002, Duclos-Cianci2010, Duclos-Cianci2010a, Wang2010, Wang2010a, Duclos-Cianci2013, Bravyi2011, Bombin2012, Wootton2012, Anwar2014, Watson2014, Hutter2014a, Bravyi2014, Wootton2015, Andrist2015}.
%~\cite{Dennis2002, Duclos-Cianci2010, Bravyi2011, Bravyi2014}.
While decoding algorithms are generally designed to run efficiently, simulating the physical error correction procedure also requires simulating the dynamics of the noise and recovery processes. 
Systems of abelian anyons are well suited to studying quantum error correction because (at the RG fixed point) these processes can be efficiently simulated numerically, allowing lattice 
simulations of decoding with over 1 million sites~\cite{Duclos-Cianci2010}. 
Standard algorithms are specifically tailored to exploit the abelian nature of these particles, and in particular the fact that abelian anyons cannot be used for quantum computation.

@@

Here we consider a full simulation of quantum error correction in a two-dimensional lattice 
system with Fibonacci anyons, a class of nonabelian anyons that are universal for quantum 
computation~\cite{Freedman2002, Nayak2008}. Fibonacci anyons are experimentally motivated as the 
expected excitations of the $\nu=\frac{12}{5}$ fractional quantum Hall 
states~\cite{Slingerland2001}, and can also be realized in several spin 
models~\cite{Levin2005, Bonesteel2012, Kapit2013, Palumbo2014} and composite 
heterostructures~\cite{Mong2014}.

@@

@@

\stf{Is there a possibility of using MPS like Rob Pfeiffer to improve the simulations or 
include interactions?}~\cite{Pfeifer2010, Pfeifer2012}

\dude{Our simulation is effectively doing a contraction of a 2d peps-like network. So
yeah, maybe there are connections.}

@@

We thank A.\ Doherty and R.\ Pfeifer for discussions. 
This work was supported by the ARC via EQuS project number CE11001013, by the US Army Research Office grant numbers W911NF-14-1-0098 and W911NF-14-1-0103, ERC grant QFTCMPS and by the cluster of excellence EXC 201 Quantum Engineering and Space-Time Research. STF also acknowledges support from an ARC Future Fellowship FT130101744.
